### 2025-11-12 decision-making.qmd
20250903 - updates to address updated data files
20251112 - typos

We created the `projectionsWithActuals_seasonal.RData` object in @sec-predictionAccuracyPrepareDataSeasonal; we created the `projectionsWithActuals_weekly.RData` object in @sec-predictionAccuracyPrepareDataWeekly.

@@

The relatively modest accuracy of the projections by so-called fantasy "experts" and of their average of their projections could occur for a number of reasons.
One possibility is that the level of expertise of the "expert" forecasters in terms of being able to provide accurate forecasts is not strong.
That is, because football performance and injuries are so challenging to predict, individual forecasters' projections may not be particularly close to the truth.

@@

Each of these possibilities is likely true to some degree.
First, individuals' predictions are unlikely to be highly accurate consistently.
Second, there are many players who are systematically *over*predicted (e.g., due to their injury) or *under*predicted (e.g., due to their becoming the starter after a teammate becomes injured, is traded, etc.)—an example of [overextremity miscalibration](#sec-calibration).
In general, it is likely for players who are projected to score more points to be overpredicted and for players who are projected to score fewer points to be underpredicted, as described in @sec-accuracyProjections.
Third, the experts may interact and discuss with one another.
Interaction and discussion among experts may lead them to follow the herd and conform their projections to what each other predict.
This has been terms informational influence and may reflect the [anchoring and adjustment heuristic](#sec-heuristicsAnchoringAdjustment) [@Larrick2024].
In any case, they are able to see each other's projections and make change their projections, accordingly.

@@

Sports analyst @Tanney2021 encourages people to consider framing your decisions as bets, to use a probabilistic mindset (rather than a determininistic mindset).
Even if something is perceived as likely to occur with 100% certainty, it is not guaranteed that such an event will occur.
Other things can and do happen, and [overconfidence](#sec-cognitiveBiasesOverconfidence) can get in the way of effective decision-making.

@@

People may make a sports bet because they believe strongly in their predicted outcome and that they "cannot be wrong".
However, the initial Vegas lines represent the aggregation of a massive amount of information (by professional oddsmakers), including team power rankings, injuries, weather, home/away, coaching, matchups, historical betting behavior, advanced statistics, and proprietary algorithms.
Leading up to the game, the Vegas lines are adjusted to balance the total money on each side of the bet (so that the sportsbook can be ensured it makes money).
In this way, the Vegas lines are based on the bets of many, many people and many dollars and reflect market wisdom.
That is, by making a sports bet, you are implicitly claiming that you have information, insight, or a model that makes you better than the market consensus at predicting the outcome—even though half of the dollars bet will lose.
Betting markets, similar to the stock market (as described in @sec-wisdomOfTheCrowd), are often highly efficient and beating them consistently is extremely challenging.
Moreover, people tend to show [confirmation bias](#sec-cognitiveBiasesConfirmation), such that they tend to remember their predictive successes and to forget their failures.
So, when they they get a bet correct, they may be more likely to bet again in the future, especially because sports betting involves [intermittent reinforcement](#sec-reinforcementIntermittent) (in particular, variable ratio reinforcement), which can make it highly addictive.

@@

In general, when you want to train a new behavior, continuous reinforcement is the fastest way to do so.
Continuous reinforcement means rewarding the behavior each time that it occurs.
For instance, a parent might use continuous reinforcement to train their child to put their toys away—the parent may praise them or give them physical affection each time they put their toys away.
However, continuous reinforcement is susceptible to extinction.
Extinction means cessation of the target behavior (putting the toys away, in this example).
That is, for a previously continusouly reinforced behavior, if the behavior stops being rewarded, the person is less likely to continue the behavior.
Thus, to make a behavior less susceptible to extinction, [intermittent reinforcement](#sec-reinforcementIntermittent) may be used.

>>>>

We created the `projectionsWithActuals_seasonal.RData` object in @sec-predictionAccuracyPrepareDataSeasonal; we created the `projectionsWithActuals_weekly.RData` object in @sec-predictionAccuracyPrepareDataWeekly.
We created the `player_stats_weekly.RData` and `player_stats_seasonal.RData` objects in @sec-calculatePlayerAge.

@@

The relatively modest accuracy of the projections by so-called fantasy "experts" and of the average of their projections could occur for a number of reasons.
One possibility is that the level of expertise of the "expert" forecasters in terms of being able to provide accurate forecasts is not strong.
That is, because football performance and injuries are so challenging to predict, individual forecasters' projections may not be particularly close to the truth.

@@

Each of these possibilities is likely true to some degree.
First, individuals' predictions are unlikely to be highly accurate consistently.
Second, there are many players who are systematically *over*predicted (e.g., due to their injury) or *under*predicted (e.g., due to their becoming the starter after a teammate becomes injured, is traded, etc.)—an example of [overextremity miscalibration](#sec-calibration).
In general, it is likely for players who are projected to score more points to be overpredicted and for players who are projected to score fewer points to be underpredicted, as described in @sec-accuracyProjections.
Third, the experts may interact and discuss with one another.
Interaction and discussion among experts may lead them to follow the herd and conform their projections to what each other predict.
This has been termed "informational influence" and may reflect the [anchoring and adjustment heuristic](#sec-heuristicsAnchoringAdjustment) [@Larrick2024].
In any case, they are able to see each other's projections and make change their projections, accordingly.

@@

Sports analyst @Tanney2021 encourages people to consider framing your decisions as bets, to use a probabilistic mindset (rather than a deterministic mindset).
Even if something is perceived as likely to occur with 100% certainty, it is not guaranteed that such an event will occur.
Other things can and do happen, and [overconfidence](#sec-cognitiveBiasesOverconfidence) can get in the way of effective decision-making.

@@

People may make a sports bet because they believe strongly in their predicted outcome and that they "cannot be wrong".
However, the initial Vegas lines represent the aggregation of a massive amount of information (by professional oddsmakers), including team power rankings, injuries, weather, home/away, coaching, matchups, historical betting behavior, advanced statistics, and proprietary algorithms.
Leading up to the game, the Vegas lines are adjusted to balance the total money on each side of the bet (so that the sportsbook can be ensured it makes money).
In this way, the Vegas lines are based on the bets of many, many people and many dollars and reflect market wisdom.
That is, by making a sports bet, you are implicitly claiming that you have information, insight, or a model that makes you better than the market consensus at predicting the outcome—even though half of the dollars bet will lose.
Betting markets, similar to the stock market (as described in @sec-wisdomOfTheCrowd), are often highly efficient and beating them consistently is extremely challenging.
Moreover, people tend to show [confirmation bias](#sec-cognitiveBiasesConfirmation), such that they tend to remember their predictive successes and to forget their failures.
So, when they get a bet correct, they may be more likely to bet again in the future, especially because sports betting involves [intermittent reinforcement](#sec-reinforcementIntermittent) (in particular, variable ratio reinforcement), which can make it highly addictive.

@@

In general, when you want to train a new behavior, continuous reinforcement is the fastest way to do so.
Continuous reinforcement means rewarding the behavior each time that it occurs.
For instance, a parent might use continuous reinforcement to train their child to put their toys away—the parent may praise them or give them physical affection each time they put their toys away.
However, continuous reinforcement is susceptible to extinction.
Extinction means cessation of the target behavior (putting the toys away, in this example).
That is, for a previously continuously reinforced behavior, if the behavior stops being rewarded, the person is less likely to continue the behavior.
Thus, to make a behavior less susceptible to extinction, [intermittent reinforcement](#sec-reinforcementIntermittent) may be used.
